Hazardous asteroids forecast via Markov random fields

### 2021-12-12 Presentation/Presentation.tex
PUB

\usepackage[bookmarks]{hyperref}
\usepackage[backend=bibtex]{biblatex}
\addbibresource{bibliography.bib}
\newtheorem*{remark}{Remark}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{Pic/Intro.jpg}}
\title[Probabilistic modeling project]{Hazardous asteroids forecast via Markov random fields}
\subtitle{Project for the exam: Probabilistic Modelling (DSE)}
\author{Marzio De Corato}
\date{\today}

@@

\begin{frame}{Celestial mechanics \cite{murray1999solar}: equations of motion}
\begin{center}
With polar coordinates  $\hat{\textbf{r}}$ and $\hat{\boldsymbol{\theta}}$
\end{center}
\begin{equation}
\textbf{r}=r\hat{\textbf{r}}
\end{equation}

@@

@@

\begin{remark}[Scientific method - Einstein definition]
Science uses the totality of the primary concepts, i.e., concepts directly connected with sense experiences, and propositions connecting them. Such a state of affairs cannot, however, satisfy a spirit which is really scientifically minded; because the totality of concepts and relations obtained in this manner is utterly lacking in logical unity. In order to supplement this deficiency, one invents a system poorer in concepts and relations, a system retaining the primary concepts and relations of the \textit{first layer} as logically derived concepts and relations. This new \textit{secondary system} pays for its higher logical unity by having elementary concepts (concepts of the second layer), which are no longer directly connected with complexes of sense experiences \cite{physics-reality}
\end{remark}

\begin{remark}[Scientific method - Einstein definition ]
The essential thing is the aim to represent the multitude of concepts and propositions, close to experience, as propositions, logically deduced from a basis, as narrow as possible, of fundamental concepts and fundamental relations which themselves can be chosen freely (axioms) \cite{physics-reality}
\end{remark}

@@

\begin{frame}{Graphical models}
\begin{theorem}[Hammersley-Clifford]
A positive distribution p(\textbf{y})>0 satisfies the CI properties of an indirect graph G iif p can be represented as a product of factor, one per maximal clique,  i.e.
\begin{equation}
p(\textbf{y}|\theta)= \dfrac{1}{Z(\theta)}\prod_{c \in C }\psi_{c}(\textbf{y}_{c}|\theta_{c})
\end{equation}
where C is the set of all the (maximal) cliques of G,  and Z($\theta$) is the partition function given by 
\begin{equation}
Z(\theta):= \sum_{y}\prod_{c\in C}\psi_{c}(\textbf{y}_{c}|\theta_{c})
\end{equation}
Note that this partition function is what ensures the overall distribution sums to 1
\end{theorem}
\end{frame}

@@

@@

\begin{frame}{mmod - Homogeneus Mixed Interaction (HMI) models}
\begin{equation}
f(\textbf{i},\textbf{y})= p(\textbf{i})(2\pi)^{-q/2}det(\Sigma)^{-1/2}\exp\left[-\dfrac{1}{2}\left(\textbf{y}-\mu(\textbf{i})\right)^{T}\Sigma^{-1}\left(\textbf{y}-\mu(\textbf{i})\right)\right]
\label{gaussMix}
\end{equation}

@@

\end{frame}
\begin{frame}{mmod}
\begin{center}
Decomposable model
\end{center}
\begin{equation}
\hat{f}(x)=\prod_{j=1}^{k}\dfrac{\hat{f}_{C_{j}}(x_{C}_{j})}{\hat{f}_{S_{j}}(x_{S}_{j})} 
\end{equation}
Likehood 
\begin{equation}
\hat{\mathcal{L}_{s}}=\sum_{i}n(i)log\left(\dfrac{n(i)}{N}\right)-Nq\dfrac{log(2pi)}{2}-N\dfrac{log(det(S))}{2}-N\frac{q}{2}
\end{equation}

>>>>

\usepackage[bookmarks]{hyperref}
\usepackage[backend=bibtex]{biblatex}
\usepackage{braket}
\addbibresource{bibliography.bib}
\newtheorem*{remark}{Remark}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{Pic/Intro.jpg}}
\title[Probabilistic modeling project]{Hazardous asteroids forecast via Markov random fields}
\subtitle{Project for the exam: Probabilistic Modelling (DSE)}
\author{Marzio De Corato}
\date{\today}

@@

\begin{frame}{Celestial mechanics \cite{murray1999solar}: equations of motion}
\begin{center}
Integrating $\textbf{r} \times \ddot{ \textbf{r}}=0$
\end{center}

\begin{equation}
\textbf{r} \times \dot{ \textbf{r}}=\textbf{h} 
\end{equation}

\begin{center}
Where \textbf{h}  is a constant of Integration. Using the polar coordinates  $\hat{\textbf{r}}$ and $\hat{\boldsymbol{\theta}}$
\end{center}

\begin{equation}
\textbf{r}=r\hat{\textbf{r}}
\end{equation}

@@

\begin{frame}{Celestial mechanics \cite{murray1999solar}: Magnitude}
\begin{equation}
\Phi=\frac{L}{4\pi d^{2}}
\end{equation}
\begin{equation}
m=-2.5\log_{10}\Phi+C
\end{equation}
\begin{equation}
m_{1}-m_{2}=-2.5\log_{10}\frac{\Phi_{1}}{\Phi_{2}}
\end{equation}
\begin{equation}
M-m=-2.5\log_{10}\frac{\Phi\cdot d^{2} }{\Phi\cdot 10^{2}}
\end{equation}
\begin{equation}
M=m+5-5log_{10}d
\end{equation}
Where $\Phi$ is the flux for a sphere of radius r, m the relative magnitude and M the \textcolor{red}{Absolute magnitude}
\end{frame}

@@

\begin{remark}[Scientific method - Einstein definition]
Science uses the totality of the primary concepts, i.e., concepts directly connected with sense experiences, and propositions connecting them. Such a state of affairs cannot, however, satisfy a spirit which is really scientifically minded;\textbf{ because the totality of concepts and relations obtained in this manner is utterly lacking in logical unity}. In order to supplement this deficiency, one invents a system poorer in concepts and relations, a system retaining the primary concepts and relations of the \textit{first layer} as logically derived concepts and relations. \textbf{This new \textit{secondary system} pays for its higher logical unity by having elementary concepts (concepts of the second layer), which are no longer directly connected with complexes of sense experiences} \cite{physics-reality}
\end{remark}

@@

\begin{frame}{Graphical models}
\begin{theorem}[Hammersley-Clifford]
A positive distribution $p(\textbf{y})>0$ satisfies the CI properties of an indirect graph G iif p can be represented as a product of factor, one per maximal clique,  i.e.
\begin{equation}
p(\textbf{y}|\theta)= \dfrac{1}{Z(\theta)}\prod_{c \in C }\psi_{c}(\textbf{y}_{c}|\theta_{c})
\end{equation}
where C is the set of all the (maximal) cliques of G,  and Z($\theta$) is the partition function given by 
\begin{equation}
Z(\theta):= \sum_{y}\prod_{c\in C}\psi_{c}(\textbf{y}_{c}|\theta_{c})
\end{equation}
Note that this partition function is what ensures the overall distribution sums to 1
\end{theorem}
\end{frame}

@@

\begin{frame}{Gaussian graphical model}
\begin{equation}
f_{Y}(y)=\dfrac{det(K)^{1/2}}{(2\pi)^{d/2}}\exp\left(-\dfrac{1}{2}\textbf{y}^{T}K\textbf{y}\right)
\end{equation}

\begin{equation}
K=\Sigma^{-1}
\end{equation}

@@

\begin{frame}{mmod - Homogeneus Mixed Interaction (HMI) models}
\begin{center}
N observation, d discrete variables and q continuous variable. The observation has the form $x=(i,y)=(i_{1},...,i_{d},y_{1}...y_{q})$. The probability of discrete variables falling in the cell i is denoted as $p(i)$. The conditional distribution of continous variables to fall in the cell $i$ is given by the multivariate gaussian $\mathcal{N}(\mu(i),\Sigma)$
\end{center}

\begin{equation}
f(\textbf{i},\textbf{y})= p(\textbf{i})(2\pi)^{-q/2}det(\Sigma)^{-1/2}\exp\left[-\dfrac{1}{2}\left(\textbf{y}-\mu(\textbf{i})\right)^{T}\Sigma^{-1}\left(\textbf{y}-\mu(\textbf{i})\right)\right]
\label{gaussMix}
\end{equation}

@@

\end{frame}
\begin{frame}{mmod}
\begin{center}
Decomposable model
\end{center}
\begin{equation}
\hat{f}(x)=\prod_{j=1}^{k}\dfrac{\hat{f}_{C_{j}}(x_{C}_{j})}{\hat{f}_{S_{j}}(x_{S}_{j})} 
\end{equation}
Maximized likehood
\begin{equation}
\hat{\mathcal{L}_{s}}=\sum_{i}n(i)log\left(\dfrac{n(i)}{N}\right)-Nq\dfrac{log(2\pi)}{2}-N\dfrac{log(det(S))}{2}-N\frac{q}{2}
\end{equation}

\begin{equation}
\hat{p}(\textbf{i})=n(\textbf{i})/N \quad \hat{\mu}(\textbf{i})=\bar{y}(\textbf{i})\quad\hat{\Sigma}=S=\sum_{i}n(i)S_{i}/N
\end{equation}
